Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-09-27 Manuscript.Rmd
noise effects draft ready

### *Potential Noise Effects*

@@

# plot
# let's put everyone's data together first
allVariance <- do.call(rbind, Variance)

@@

# There is a strong linear relationship between the Fiedler vector and the R-squared values
# These match perfectly on the brain when inspected qualitatively
v1 <- ggplot(data = allVariance) + 
  geom_smooth(aes(FiedlerVec, Rsquared, group = SubjID), fill = "gray70", color = "gray30", show.legend = F) + 
  labs(x = "Fiedler Vector Values", y = "R-Squared Values") + 
  theme_classic()

@@

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001) (Figure 14). These differences had moderately high effect sizes (mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole.

@@

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while the highest $R^2$ values are indeed placed dorsally, there is no clear relationship with the ventral-dorsal axis. However, the coefficient of variation derived from minimally preprocessed data does show a trend (namely, the more ventral, the lower the SNR). This is notable because the R-squared values are estimated from data with added noise reduction steps. These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that our stability results could be due mostly to highly structured mPFC noise.

``` {r Noise Effects: R-sqd vs tSNR (Fig. 15), fig.align="center", fig.width=5, fig.height=3, fig.cap = "Figure 15. Vertex tSNR values across all subjects are correlated with their position along the z-axis in mPFC, with vmPFC showing the least tSNR (left). However, R-square values do not show this strong relationship (right), suggesting that the coupling of each vertex is not necessarily influenced by signal noise.", echo=FALSE, message = FALSE, cache = TRUE}
# grid.arrange(v6, v7, ncol = 2)
# par(mfrow = c(1,2))
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$Rsquared, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "R-squared Value",
#               ylab = "Z-Coordinate")
# 
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$varCoeff, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "Coefficient of Variation (SNR)",
#               ylab = "")

tempVariance <- allVariance[allVariance$Region == "mPFC", ]
s1 <- ggplot(data = allVariance, aes(Rsquared, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "R-squared value", y = "Z-Coordinate") +
  theme_classic()

s2 <- ggplot(data = allVariance, aes(varCoeff, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "tSNR", y = "Z-Coordinate") +
  theme_classic()

@@

It's also worth mentioning that PCC is largely covered by DMN. Then why does this area overlap with SV? I think the Mueller paper mentions that PCC is an important hub of the brain, and perhaps subserves many cognitive functions.

@@

>>>>

### *Potential Noise Effects (this part is challenging)*

@@

# let's put everyone's data together first
allVariance <- do.call(rbind, Variance)

@@

# Create a new data frame with values restricted to mPFC (for plotting and further stats)
mPFCVariance <- allVariance[allVariance$Region == "mPFC", ]

# Get the correlations between Z-axis, R2, and tSNR for every subject (to take mean and SD in text)
# I just came up with this, so the code should be made more parsimonious in the future (like, why bind and then split...?)
tempSubj <- split(mPFCVariance, f = mPFCVariance$SubjID)
noiseCors <- data.frame(R2vtSNR = sapply(tempSubj, function(data) {cor(data$Rsquared, data$varCoeff)}),
                        ZvtSNR = sapply(tempSubj, function(data) {cor(data$tempZ, data$varCoeff)}),
                        ZvR2 = sapply(tempSubj, function(data) {cor(data$tempZ, data$Rsquared)}))
rm(tempSubj)

# plot
# There is a strong linear relationship between the Fiedler vector and the R-squared values
# These match perfectly on the brain when inspected qualitatively
v1 <- ggplot(data = allVariance) + 
  geom_smooth(aes(FiedlerVec, Rsquared, group = SubjID), fill = "gray70", color = "gray30", show.legend = F) + 
  labs(x = "Fiedler Vector Values", y = "R-Squared Values") + 
  theme_classic()

@@

@@

To further interrogate the potential effect of noise in our results, we computed the variance explained by activity from each DMN vertex on the mean time series of the DMN community. If the signal quality issues persisted after preprocessing the data, we expected that noisy vertices would have the least explicative power on the activity of the DMN community as a whole. This also implied lower $R^2$ values for mPFC. Indeed, we found that for each subject, $R^2$ values tended to be higher for PCC (`r prcntSig` percent of Wilcoxon ranksum tests across subjects had *p* values < 0.001) (Figure 14). These differences had moderately high effect sizes (mean Cohen's D = `r round(mean(Rsq_PCCvsPFC_ES), digits = 2)`). These results suggest that, overall, vertices in mPFC tended to be less in synchrony with the activity of the DMN community as a whole. Furthermore, the degree tSNR showed a moderate correlation with $R^2$ values across participants (mean *r* = `r round(mean(noiseCors$R2vtSNR), digits = 2)`; SD = `r round(sd(noiseCors$R2vtSNR), digits = 2)`), meaning that noisier vertices were less in synchrony with the general activity of the DMN.

Given this trend, we asked whether there is a "gradient of noise" along the z-axis of mPFC, such that more ventral vertices would both be noisier and explain the least amount of DMN variance. Here we found that while $R^2$ are somewhat correlated with dorsal-ventral placement across subjects (mean *r* = `r round(mean(noiseCors$ZvR2), digits = 2)`; SD = `r round(sd(noiseCors$ZvR2), digits = 2)`), there is no clear relationship with the z-axis across subjects (Figure 15). Specifically, low $R^2$ are distributed evenly along this axis, denoting a degree of orthogonality between these features. Conversely, tSNR showed a strong linear relationship with the position of the vertex along this axis (mean *r* = `r round(mean(noiseCors$ZvtSNR), digits = 2)`; SD = `r round(sd(noiseCors$ZvtSNR), digits = 2)`). These results contextualize the degree of synchrony of mPFC vertices in terms of potential noise. While still an important factor, it is unlikely that the degree of within-subject pattern stability observed here could be due mostly to highly structured mPFC noise.

``` {r Noise Effects: R-sqd vs tSNR (Fig. 15), fig.align="center", fig.width=5, fig.height=3, fig.cap = "Figure 15. Vertex tSNR values across all subjects are correlated with their position along the z-axis in mPFC, with vmPFC showing the lowest tSNR (left). However, R-square values do not show this strong relationship (right), suggesting that the coupling of each vertex is not necessarily influenced by signal noise.", echo=FALSE, message = FALSE, cache = TRUE}
# grid.arrange(v6, v7, ncol = 2)
# par(mfrow = c(1,2))
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$Rsquared, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "R-squared Value",
#               ylab = "Z-Coordinate")
# 
# smoothScatter(allVariance[allVariance$Region == "mPFC",]$varCoeff, allVariance[allVariance$Region == "mPFC",]$tempZ,
#               xlab = "Coefficient of Variation (SNR)",
#               ylab = "")

# ALTERNATIVE IDEA FOR A FIGURE: BRAIN MAPS WITH THESE VALUES. THAT SAID, R2 MAPS LOOK THE SAME AS FV ONES, SO...
s1 <- ggplot(data = mPFCVariance, aes(Rsquared, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "R-squared value", y = "Z-Coordinate") +
  theme_classic()

s2 <- ggplot(data = mPFCVariance, aes(varCoeff, tempZ)) +
  geom_bin2d(bins = 250) +
  scale_fill_continuous(low = "grey30", high = "grey70") +
  geom_density2d(color = Cols[2]) +
  labs(x = "tSNR", y = "Z-Coordinate") +
  theme_classic()

@@

It's also worth mentioning that PCC is largely covered by DMN. Then why does this area overlap with SV? I think the Mueller paper mentions that PCC is an important hub of the brain, and perhaps subserves many cognitive functions (recent Wash U paper on this too).

@@

In terms of noise, there are a handful of things to keep in mind. First is that yes, mPFC is noisier overall, but that did not preclude the relatively high levels of agreement across days and sessions per subject, even though ARI heavily penalizes discrepancies (less than linearly). Second, low $R^2$ values were found all throughout the z-axis, meaning that variance explained is not strictly determined by the cardinal position of the vertex (and while the correlation was significant due to the sheer number of items, the coefficient was relatively low). Right now this might be ok as a way to bypass noise comments, but it would also be interesting to take advantage of the Fiedler vector here. Since it's linearly related to $R^2$ per subject, perhaps finding a way to threshold these values would provide cleaner and even more consistent maps.
